---
title: Invasion Dynamics of a Brittle Star at Different Spatial Scales across the
  West Atlantic
author: "Jospeh Benz"
---

# Introduction

*Ophiothela* *mirabilis* is a tiny (~ 5 mm) orange brittle star natively found on Pacific tropical reefs. It lives on structures like hard corals, soft corals, and sponges, where it feeds on particulates in the water column. In the year 2000, it was found on reefs in coastal Brazil, and has been spreading upwards through the Caribbean since, recently discovered in Florida in 2020.  In this invasive range, *O.* *mirabilis* generally reproduces asexually, splitting down the middle in a process called fissiparity, allowing it to rapidly spread across the reef. However, the mechanism for this spread is largely unknown; they are negatively buoyant and very poor swimmers, and are likely taking advantage of some other method to travel the great distances between islands in the Caribbean.  The two prevailing hypothesis include *O.* *mirabilis* being transported by shipping traffic (either in bilge water or on boat hulls) or by “rafting” on patches of *Sargassum*, a floating macroalgae common to the area. 
In order to determine the invasion dynamics of this species, we examine the data at two different spatial scales: The West Atlantic Tropics and the island of St. John, USVI.  Looking at the entire west Atlantic will allow us to examine the saltatory spread (large jumps across habitat gaps), while focusing in on a single island informs the smaller scale spread between adjacent reefs after the brittle star has arrived. 

# Materials and methods

## Required packages:
```{r}
library(shiny)
library(leaflet)
library(sp)
library(spocc)
library(raster)
library(tidyverse)
library(leaflet.extras2)
library(sf)
library(rsconnect)
```
## Data Download & Processing
First, a bounding box for the larger spatial scale is created.
```{r}
Carib_extent <- extent(c(xmin=-130,xmax=-17,ymin=-26,ymax=33))
bbox1 <- bbox(Carib_extent)
```

Using the occ() function from the spocc package (), we can examine occurrence data for *Ophiothela* *mirabilis* within our bounding box, utilizing an array of online databases. In this case, species records are only used from the Global Biodiversity Information Facility (GBIF) to prevent the duplication of observations. Then, occ2df() will download and convert our occurrences into a data frame with latitude and longitude. Naming convention varies, so another column is added with mutate() to create a consistent name column.   
```{r}
OM <- occ("Ophiothela mirabilis", from = c('gbif'), has_coords=TRUE, limit=1000000, geometry=bbox1)
OMdf <- occ2df(OM) %>%
  mutate(simplename="Ophiothela mirabilis")
```
















Narrative: Clear narrative description of the data sources and methods. Includes data from at least two sources that were integrated / merged in R.

Code: The code associated with the project is well organized and easy to follow. Demonstrates mastery of R graphics and functions.

Data: The underlying data are publicly accessible via the web and downloaded/accessed within the Rmd script. If you want to use your own data, you must make it available on a website (e.g. Figshare) so that others are able to re-run your code.

You can do bullets like this:

* The first most important thing
* The second most important thing
* The third most important thing

You can do numbers like this:

1. The first most important thing
2. The second most important thing
3. The third most important thing

See [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/) for all the amazing things you can do.


Here's my first code chunk.
```{r}
1+2
```

Load any required packages in a code chunk (you may need to install some packages):

```{r, message=F, warning=F}
library(tidyverse)
library(leaflet)
library(kableExtra)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data
```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

Add any additional processing steps here.

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r}
library(shiny)
library(leaflet)
library(sp)
library(spocc)
library(raster)
library(tidyverse)
library(leaflet.extras2)
library(sf)
```
```{r}
Carib_extent <- extent(c(xmin=-130,xmax=-17,ymin=-26,ymax=33))
bbox1 <- bbox(Carib_extent)

OM <- occ("Ophiothela mirabilis", from = c('gbif','bison','inat', 'ebird', 'vertnet'), has_coords=TRUE, limit=1000000, geometry=bbox1)
OMdf <- occ2df(OM) %>%
  mutate(simplename="Ophiothela mirabilis")

OMdf$Year <- format(as.Date(OMdf$date, format="%Y/%m/%d"),"%Y")  

OMdf2 <- OMdf %>%
  filter(Year!=1971)
```
```{r cache=FALSE}
ui <- bootstrapPage(
  tags$style(type = "text/css", "html, body {width:100%;height:100%}"),
  leafletOutput("map", width = "100%", height = "100%"),
  absolutePanel(bottom = 30, right = 10,
                sliderInput("Year", "Year", 1995, 2021, value = 1995, step = 1, sep = "")
  )
)

server <- function(input, output, session) {
  
  sliderData <- reactive({
    OMdf2 %>%
      filter(Year <= input$Year)
  })
  
  output$map <- renderLeaflet({
    leaflet() %>%
      addTiles() %>%
      fitBounds(min(OMdf2$longitude), min(OMdf2$latitude),
                max(OMdf2$longitude), max(OMdf2$latitude))
  })
  
  observe({
    leafletProxy("map", data = sliderData()) %>%
      clearMarkers() %>%
      addTiles() %>%
      addMarkers(lng = ~longitude, lat = ~latitude)
  })
}

shinyApp(ui = ui, server = server)
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner


